\propertitle

### 2019-12-06 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-05T23:42:10+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The $T= \infty$ approximation can break the method for specific
distributions of sample frequencies, if too many moments are used. For
example take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
activities $\ya=0,1,2$; that is, the sample is silent half of the time and
completely active half of the time. The first two normalized factorial
moments have both value $1/2$, and together they uniquely determine $\yf$.
If we now take $N=3$, it is possible to show that
\eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two moments, has
no solution. The simple reason is clear from a sampling perspective: if
there is zero probability of sampling one active neuron, then there must be
zero probability of sampling two active ones. Without the $T=\infty$
approximation this doesn't happen, because the conditional frequency of the
activity levels and the sampling conditional probability are kept distinct.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-06T11:13:57+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The $T= \infty$ approximation can break the method for specific
distributions of sample frequencies, if too many moments are used. For
example take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
activities $\ya=0,1,2$; that is, the sample is silent half of the time and
completely active half of the time. The first two normalized factorial
moments have both value $0.5$, and together they uniquely determine $\yf$.
If we now take $N=3$, it can be proved that
\eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two moments, has
no solution. The reason is clear from a sampling perspective: if there is
zero probability of sampling one active neuron, then the larger population
cannot have any active neurons, and there must be zero probability of
sampling two active ones. But this reasoning holds only for each single
time bin, not for the averages over time bins. The $T=\infty$ approximation
identifies the two instead. Outside this approximation there are no
contradictions because the conditional frequencies and conditional
probabilities are kept distinct.
